\chapter{Future Work: Application}
\label{chapter:application}

The goal of this thesis is to investigate whether the evolution of mimicry can be used in a problem solving scenario similar to evolutionary programming or genetic algorithm as it is used to optimize solutions using mathematical functions. Predator-prey co-evolution has also been used for solving interesting problems \cite{hillis1990}, giving better result than conventional methodologies. 

The biggest challenge faced while searching for the appropriate problem to be optimized is the idea of associating palatability with CA pattern with which the prey species has been represented. If we consider the CA pattern to be a certain solution to a problem then the set of 2D CA pattern among which the prey species vary could be considered as the solution set of the problem. And if we consider that the predator species are responsible for selecting the appropriate pattern based on their palatability, then we can consider palatability as the criteria for selection of the pattern (or the evaluation/fitness function for the case of Evolutionary Programming/Genetic Algorithm). If the predator associates the CA pattern with palatability then it does not make any sense of having a mimic in the solution space as mimics will have the same pattern with opposite palatability. So using the evolution of mimicry to solve an optimization problem is futile if we try conventional problems which already can be solved with the help of evolutionary programming or genetic algorithm. 

So the kind of problems we should be looking for applying the evolution of mimicry, are the ones which cannot be solved with conventional evolutionary methods. We should be looking for a problem where the idea of deception can be useful in term of solving it. An interesting analogy is with anti-virus and virus software. If we consider anti-virus to be predator and virus (mimic) to be prey species, then the objective of a virus would be to look like a regular virus free program (model) to get pass an anti-virus software. In this way by mimicking a regular program a virus can get pass any anti-virus software available in the market. To successfully implement this software we need to create a virus program which can evolve itself to pretend like a regular program (virus free). The environment will be the operating system itself where all the programs reside. It is worth noting the fact that this objective is malicious for the computer industry, which makes it less significant to pursue. 